---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# multivaR

<!-- badges: start -->
<!-- badges: end -->

The goal of multivaR is to organize and simplify the process of creating a multivariate analysis. This package can format data as well as caluclate anomalies from raw data, once data is formatted the package can be used to run Principal Components Analysis on multiple variables. 

## Installation

You can install  the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Echisholm21/multivaR")
```

multivaR is designed to work with other packages developed by the Data Access team at Bedford Institute of Oceanogrpahy (BIO). Specifically, this package is designed to work with data which is supplied through the `azmpdata` package. This supplementary package can be installed by using:

``` r
devtools::install_github("casaultb/azmpdata")
```



## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(multivaR)
## basic example code
```


