---
title: "draft"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{draft}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(multivaR)
library(dplyr)
library(data.table)
```


This is a rough draft vignette of how to run a typical workflow using multivaR


Step one is to assemble all the data for different variables into a single dataframe

```{r}

# step 1: assemble data

# search for azmpdata
# if azmpdata is not downloaded:
# devtools::install_github("casaultb/azmpdata")
library(azmpdata)
query(search_key = '')

# load in desired data frames
data("nitrate_inventory_timeseries_hl2")
data("chlorophyll_inventory_timeseries_hl2")

# collate data frames

datfr <- right_join(chlorophyll_inventory_timeseries_hl2, nitrate_inventory_timeseries_hl2)

```

If anomalies need to be calculated, multivaR has options for calculating monthly or annual anomalies with or without normalization.

```{r}
# monthly anomalies must be caluclated before annual anomalies (annual anomaly function simply combines monthly anomalies)

danomalies <- monthlyAnomaly(datfr, climatologyYears = c(1999, 2010), var = 'chl_0_100')

# annual anomalies function creates new data frame with just year and variable
chl_annual_anomalies <- annualAnomaly(danomalies, anomaly = 'chl_0_100_anomaly')

```
A note on caluclatinng anomalies, reference periods are used to create climatologies to calculate anomalies in this method. If you are unsure of a proper reference period for your data, follow these rules:

* 


Once data is assembled, PCA can be run


```{r}
PCA(datfr)

```




